### 2025-02-21 code/eda-code/eda.qmd
Part 2 of MADA Project

---
title: "Exploratory analysis"
---

We start of by summarizing all the information

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyr)
library(knitr)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.
First, we summarize all our information using a table.

```{r}
library(dplyr)
library(tidyr)
library(kableExtra)

# Create a table for Age summary with Mean and SD
age_summary <- data.frame(
    Category = "Age",
    Value = paste(round(mean(mydata$patient_age, na.rm = TRUE), 2), "±", round(sd(mydata$patient_age, na.rm = TRUE), 2)),
    Count = NA,
    Percent = NA
)

# Sex Summary
sex_summary <- mydata %>%
    group_by(sex) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = sex) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Drug Summary
drug_summary <- mydata %>%
    group_by(suspect_product_active_ingredients) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = suspect_product_active_ingredients) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Outcome Summary
outcome_summary <- mydata %>%
    group_by(serious) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = serious) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reporter Type Summary
reporter_summary <- mydata %>%
    group_by(reporter_type) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = reporter_type) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Country Summary
country_summary <- mydata %>%
    group_by(country_where_event_occurred) %>%
    summarise(Count = n()) %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    rename(Category = country_where_event_occurred) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Reaction Type Summary
# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

summary_data <- mydata %>%
    select(all_of(organ_systems)) %>%
    pivot_longer(cols = everything(), values_to = "organ_system") %>%
    filter(!is.na(organ_system)) %>%
    group_by(organ_system) %>%
    summarise(
        Count = n(),
        Percent = round((n() / nrow(mydata)) * 100, 1)
    ) %>%
    rename(Category = organ_system) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Combine all summaries into one table
final_summary_table <- bind_rows(
    data.frame(Category = "Age", Value = age_summary$Value, stringsAsFactors = FALSE),
    data.frame(Category = "Sex", Value = "", stringsAsFactors = FALSE), # Subheading for Sex
    sex_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Sex
    data.frame(Category = "Immune Checkpoint Inhibitors", Value = "", stringsAsFactors = FALSE), # Subheading for Drugs
    drug_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Drugs
    data.frame(Category = "Outcome", Value = "", stringsAsFactors = FALSE), # Subheading for Outcome
    outcome_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Outcome
    data.frame(Category = "Reporter Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reporter Type
    reporter_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Reporter Type
    data.frame(Category = "Country", Value = "", stringsAsFactors = FALSE), # Subheading for Country
    country_summary %>% select(Category, Value),
    data.frame(Category = "", Value = "", stringsAsFactors = FALSE), # Blank row after Country
    data.frame(Category = "Reaction Type", Value = "", stringsAsFactors = FALSE), # Subheading for Reaction Type
    summary_data %>% select(Category, Value)
)

# Create a publication-quality table using kableExtra
final_summary_table %>%
    kable("html", caption = "Table 1: Descriptive Statistics", col.names = c("Category", "Value")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
    column_spec(1, bold = TRUE) %>%
    row_spec(which(final_summary_table$Category %in% c("Age", "Sex", "Immune Checkpoint Inhibitors", "Outcome", "Reporter Type", "Country", "Reaction Type")), bold = TRUE, background = "#f7f7f7") %>%
    footnote(general = "This table summarizes the patient data by various categories.")

```

We save the table.

```{r}
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(final_summary_table, file = summarytable_file)
```

# Data exploration through figures

Bar graph for outcomes by drugs.

```{r}
library(ggplot2)

# Define custom colors for the 'serious' variable
custom_colors <- c(
    "Serious" = "#d7191c",  # Red for serious
    "Non-Serious" = "#2c7bb6"    # Blue for non-serious
)

# Create the plot with custom colors for 'serious'
p1 <- mydata %>%
    ggplot(aes(x = suspect_product_active_ingredients, fill = serious)) +
    geom_bar(position = "stack") + # Use "stack" for stacked bars
    scale_fill_manual(values = custom_colors) + # Apply custom colors
    labs(
        title = "Bar Plot of Outcomes of Immune Checkpoint Inhibitors",
        x = "Immune Checkpoint Inhibitors",
        y = "Count",
        fill = "Outcomes"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

# Print the plot
print(p1)

figure_file = here("results", "figures","figure1.png")
ggsave(filename = figure_file, plot=p1) 
```

Now let's look at the bar graph by type of organ system.

```{r}
p2 <- summary_data %>%
    ggplot(aes(x = reorder(Category, Count), y = Count)) + # Reorder by Count ascending
    geom_col(fill = "#1f77b4") +
    labs(
        title = "Bar Plot of Reaction Types",
        x = "Organ System",
        y = "Count",
        caption = "Percentages are shown in parentheses"
    ) +
    theme_minimal() +
    geom_text(aes(label = Value), hjust = -0.1, size = 3) + # Adjust label position
    coord_flip() # Flip the axes

# Print the plot
print(p2)

figure_file = here("results", "figures","figure2.png")
ggsave(filename = figure_file, plot=p2) 
```