### 2025-03-14 code/analysis-code/statistical-analysis.qmd
Part 3 of MADA Project

---
title: "statistical-analysis"
format: html
---

This file contains details of the statistical analysis.

After the exploratory data analysis and the descriptive statistics, the next step is to run a disproportionality analysis. The first step is to assess the adverse events in comparison to the entire FAERS database.

To assess this, we need the following information: 
a = Number of irAEs with 1 ICI (e.g. atezolizumab) 
b = Number of reactions other than irAEs with atezolizumab
c = Number of irAEs with drugs other than atezolizumab 
d = Number of reactions other than irAEs with drugs other than atzeolizumab.

The same procedure would be repeated for the other 3 ICIs.

First, we load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

First we find number of irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions <- function(data, drug) {
  # Filter rows for the drug
  drug_data <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions <- c(
    drug_data$reaction_1,
    drug_data$reaction_2,
    drug_data$reaction_3,
    drug_data$reaction_4,
    drug_data$reaction_5
  )

  # Remove NA values
  all_reactions <- all_reactions[!is.na(all_reactions)]

  # Count total reactions
  total_reactions <- length(all_reactions)

  # Return total reactions
  return(total_reactions)
}

# Count total reactions for each drug
results <- sapply(drugs, function(drug) count_total_reactions(mydata, drug))

# Print the results
results_irae <- data.frame(Drug = drugs, Total_Reactions = results)
print(results_irae)
```

The `echo: false` option disables the printing of code (only output is displayed).

Now, we do the same for reactions other than irAEs. We created and saved this data when we processed it.

Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata1.rds")
#load data
mydata1 <- readRDS(data_location)
```

Now we find number of reactions other than irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions1 <- function(data, drug) {
  # Filter rows for the drug
  drug_data1 <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions1 <- c(
    drug_data1$reaction_1,
    drug_data1$reaction_2,
    drug_data1$reaction_3,
    drug_data1$reaction_4,
    drug_data1$reaction_5
  )

  # Remove NA values
  all_reactions1 <- all_reactions1[!is.na(all_reactions1)]

  # Count total reactions
  total_reactions1 <- length(all_reactions1)

  # Return total reactions
  return(total_reactions1)
}

# Count total reactions for each drug
results1 <- sapply(drugs, function(drug) count_total_reactions1(mydata1, drug))

# Print the results
results_non_irae <- data.frame(Drug = drugs, Total_Reactions = results1)
print(results_non_irae)
```
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for All irAEs
results_table_all_iraes <- data.frame()

# Loop through each drug and perform the analysis for all irAEs
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for all irAEs combined)
  a <- results_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  b <- results_non_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  c <- results_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  d <- results_non_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a / (a + b)) / (c / (c + d))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a) - (1 / (a + b)) + (1 / c) - (1 / (c + d)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a / b) / (c / d)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a) + (1 / b) + (1 / c) + (1 / d))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a * (a + b + c + d)) / ((a + c) * (a + b)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a - 1 / (a + b) - 1 / (a + c) + 1 / (a + b + c + d))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (All irAEs)
  results_table_all_iraes <- rbind(results_table_all_iraes, data.frame(
    Drug = drug_of_interest,
    Reaction = "All irAEs",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for All irAEs
print(results_table_all_iraes)
```
The next step is to assess the safety signal for individual irAEs.
The commonly reported irAEs in clinical trials are rash, pruritus, colitis, hepatitis, hyperthyroidism, hypothyroidism, and nephritis.
We conduct disproportionality analysis for each of these.

We compare each of these reactions to all other reactions.
First, we assess the number of 'rash' reactions reported with all the ICIs.

```{r}
 # Filter for the specific drugs and count 'Rash' reactions
rash_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Rash = sum(c_across(starts_with("reaction_")) == "Rash", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Rash_Count = sum(Rash))

# Print the result
print(rash_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'rash' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

rash_counts <- rash_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Rash' counts from total reactions
results_non_rash <- results_irae %>%
  mutate(
    Non_Rash_Reactions = Total_Reactions - rash_counts$Rash_Count
  )

# Print the updated results
print(results_non_rash)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_rash <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_rash$Non_Rash_Reactions
)

# Print the new dataset
print(results_other_rash)

```

We have all our numbers now.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Rash
results_table_rash <- data.frame()

# Loop through each drug and perform the analysis for Rash
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Rash)
  a_rash <- rash_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Rash_Count)
  b_rash <- results_other_rash %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_rash <- rash_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Rash_Count)) %>% pull()
  d_rash <- results_other_rash %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_rash / (a_rash + b_rash)) / (c_rash / (c_rash + d_rash))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_rash) - (1 / (a_rash + b_rash)) + (1 / c_rash) - (1 / (c_rash + d_rash)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_rash / b_rash) / (c_rash / d_rash)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_rash) + (1 / b_rash) + (1 / c_rash) + (1 / d_rash))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_rash * (a_rash + b_rash + c_rash + d_rash)) / ((a_rash + c_rash) * (a_rash + b_rash)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_rash - 1 / (a_rash + b_rash) - 1 / (a_rash + c_rash) + 1 / (a_rash + b_rash + c_rash + d_rash))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Rash)
  results_table_rash <- rbind(results_table_rash, data.frame(
    Drug = drug_of_interest,
    Reaction = "Rash",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Rash
print(results_table_rash)

```

Now, we conduct the same analysis for pruritus.
```{r}
 # Filter for the specific drugs and count 'Pruritus' reactions
pruritus_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Pruritus = sum(c_across(starts_with("reaction_")) == "Pruritus", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Pruritus_Count = sum(Pruritus))

# Print the result
print(pruritus_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'pruritus' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

pruritus_counts <- pruritus_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'pruritus' counts from total reactions
results_non_pruritus <- results_irae %>%
  mutate(
    Non_Pruritus_Reactions = Total_Reactions - pruritus_counts$Pruritus_Count
  )

# Print the updated results
print(results_non_pruritus)

# Create the new dataset by adding Non_Pruritus_Reactions to Total_Reactions
results_other_pruritus <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_pruritus$Non_Pruritus_Reactions
)

# Print the new dataset
print(results_other_pruritus)

```
We now conduct the analysis.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Pruritus
results_table_pruritus <- data.frame()

# Loop through each drug and perform the analysis for Pruritus
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Pruritus)
  a_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Pruritus_Count)
  b_pruritus <- results_other_pruritus %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Pruritus_Count)) %>% pull()
  d_pruritus <- results_other_pruritus %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_pruritus / (a_pruritus + b_pruritus)) / (c_pruritus / (c_pruritus + d_pruritus))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_pruritus) - (1 / (a_pruritus + b_pruritus)) + (1 / c_pruritus) - (1 / (c_pruritus + d_pruritus)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_pruritus / b_pruritus) / (c_pruritus / d_pruritus)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_pruritus) + (1 / b_pruritus) + (1 / c_pruritus) + (1 / d_pruritus))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_pruritus * (a_pruritus + b_pruritus + c_pruritus + d_pruritus)) / ((a_pruritus + c_pruritus) * (a_pruritus + b_pruritus)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_pruritus - 1 / (a_pruritus + b_pruritus) - 1 / (a_pruritus + c_pruritus) + 1 / (a_pruritus + b_pruritus + c_pruritus + d_pruritus))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Pruritus)
  results_table_pruritus <- rbind(results_table_pruritus, data.frame(
    Drug = drug_of_interest,
    Reaction = "Pruritus",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Pruritus
print(results_table_pruritus)

```

We repeat this analysis for hypothyroidism.
```{r}
 # Filter for the specific drugs and count 'Hypothyroidism' reactions
hypothyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hypothyroidism = sum(c_across(starts_with("reaction_")) == "Hypothyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hypothyroidism_Count = sum(Hypothyroidism))

# Print the result
print(hypothyroidism_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hypothyroidism' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hypothyroidism_counts <- hypothyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hypothyroidism' counts from total reactions
results_non_hypothyroidism <- results_irae %>%
  mutate(
    Non_Hypothyroidism_Reactions = Total_Reactions - hypothyroidism_counts$Hypothyroidism_Count
  )

# Print the updated results
print(results_non_hypothyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hypothyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hypothyroidism$Non_Hypothyroidism_Reactions
)

# Print the new dataset
print(results_other_hypothyroidism)

```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hypothyroidism
results_table_hypothyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hypothyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hypothyroidism)
  a_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hypothyroidism_Count)
  b_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hypothyroidism_Count)) %>% pull()
  d_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hypothyroidism / (a_hypothyroidism + b_hypothyroidism)) / (c_hypothyroidism / (c_hypothyroidism + d_hypothyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hypothyroidism) - (1 / (a_hypothyroidism + b_hypothyroidism)) + (1 / c_hypothyroidism) - (1 / (c_hypothyroidism + d_hypothyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hypothyroidism / b_hypothyroidism) / (c_hypothyroidism / d_hypothyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hypothyroidism) + (1 / b_hypothyroidism) + (1 / c_hypothyroidism) + (1 / d_hypothyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hypothyroidism * (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism)) / ((a_hypothyroidism + c_hypothyroidism) * (a_hypothyroidism + b_hypothyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hypothyroidism - 1 / (a_hypothyroidism + b_hypothyroidism) - 1 / (a_hypothyroidism + c_hypothyroidism) + 1 / (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hypothyroidism)
  results_table_hypothyroidism <- rbind(results_table_hypothyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hypothyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hypothyroidism
print(results_table_hypothyroidism)

```

Now, we assess the reports for hyperthyroidism:
```{r}
 # Filter for the specific drugs and count 'hyperthyroidism' reactions
hyperthyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hyperthyroidism = sum(c_across(starts_with("reaction_")) == "Hyperthyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hyperthyroidism_Count = sum(Hyperthyroidism))

# Print the result
print(hyperthyroidism_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hyperthyroidism' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hyperthyroidism_counts <- hyperthyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hyperthyroidism' counts from total reactions
results_non_hyperthyroidism <- results_irae %>%
  mutate(
    Non_Hyperthyroidism_Reactions = Total_Reactions - hyperthyroidism_counts$Hyperthyroidism_Count
  )

# Print the updated results
print(results_non_hyperthyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hyperthyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hyperthyroidism$Non_Hyperthyroidism_Reactions
)

# Print the new dataset
print(results_other_hyperthyroidism)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hyperthyroidism
results_table_hyperthyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hyperthyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hyperthyroidism)
  a_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hyperthyroidism_Count)
  b_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hyperthyroidism_Count)) %>% pull()
  d_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hyperthyroidism / (a_hyperthyroidism + b_hyperthyroidism)) / (c_hyperthyroidism / (c_hyperthyroidism + d_hyperthyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hyperthyroidism) - (1 / (a_hyperthyroidism + b_hyperthyroidism)) + (1 / c_hyperthyroidism) - (1 / (c_hyperthyroidism + d_hyperthyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hyperthyroidism / b_hyperthyroidism) / (c_hyperthyroidism / d_hyperthyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hyperthyroidism) + (1 / b_hyperthyroidism) + (1 / c_hyperthyroidism) + (1 / d_hyperthyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hyperthyroidism * (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism)) / ((a_hyperthyroidism + c_hyperthyroidism) * (a_hyperthyroidism + b_hyperthyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hyperthyroidism - 1 / (a_hyperthyroidism + b_hyperthyroidism) - 1 / (a_hyperthyroidism + c_hyperthyroidism) + 1 / (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hyperthyroidism)
  results_table_hyperthyroidism <- rbind(results_table_hyperthyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hyperthyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hyperthyroidism
print(results_table_hyperthyroidism)

```

Now, we repeat this analysis with colitis:

```{r}
 # Filter for the specific drugs and count 'colitis' reactions
colitis_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Colitis = sum(c_across(starts_with("reaction_")) == "Colitis", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Colitis_Count = sum(Colitis))

# Print the result
print(colitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'colitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

colitis_counts <- colitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'colitis' counts from total reactions
results_non_colitis <- results_irae %>%
  mutate(
    Non_Colitis_Reactions = Total_Reactions - colitis_counts$Colitis_Count
  )

# Print the updated results
print(results_non_colitis)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_colitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_colitis$Non_Colitis_Reactions
)

# Print the new dataset
print(results_other_colitis)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Colitis
results_table_colitis <- data.frame()

# Loop through each drug and perform the analysis for Colitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Colitis)
  a_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Colitis_Count)
  b_colitis <- results_other_colitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Colitis_Count)) %>% pull()
  d_colitis <- results_other_colitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_colitis / (a_colitis + b_colitis)) / (c_colitis / (c_colitis + d_colitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_colitis) - (1 / (a_colitis + b_colitis)) + (1 / c_colitis) - (1 / (c_colitis + d_colitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_colitis / b_colitis) / (c_colitis / d_colitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_colitis) + (1 / b_colitis) + (1 / c_colitis) + (1 / d_colitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_colitis * (a_colitis + b_colitis + c_colitis + d_colitis)) / ((a_colitis + c_colitis) * (a_colitis + b_colitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_colitis - 1 / (a_colitis + b_colitis) - 1 / (a_colitis + c_colitis) + 1 / (a_colitis + b_colitis + c_colitis + d_colitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Colitis)
  results_table_colitis <- rbind(results_table_colitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Colitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Colitis
print(results_table_colitis)

```
Now, we conduct the same analysis for hepatitis:

```{r}
# Filter for the specific drugs and count 'hepatitis' reactions
hepatitis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Hepatitis = sum(c_across(starts_with("reaction_")) == "Hepatitis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Hepatitis_Count = sum(Hepatitis))

# Print the result
print(hepatitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hepatitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hepatitis_counts <- hepatitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'hepatitis' counts from total reactions
results_non_hepatitis <- results_irae %>%
  mutate(
    Non_Hepatitis_Reactions = Total_Reactions - hepatitis_counts$Hepatitis_Count
  )

# Print the updated results
print(results_non_hepatitis)

# Create the new dataset by adding Non_Hepatitis_Reactions to Total_Reactions
results_other_hepatitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hepatitis$Non_Hepatitis_Reactions
)

# Print the new dataset
print(results_other_hepatitis)
```
Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hepatitis
results_table_hepatitis <- data.frame()

# Loop through each drug and perform the analysis for Hepatitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hepatitis)
  a_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hepatitis_Count)
  b_hepatitis <- results_other_hepatitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hepatitis_Count)) %>% pull()
  d_hepatitis <- results_other_hepatitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hepatitis / (a_hepatitis + b_hepatitis)) / (c_hepatitis / (c_hepatitis + d_hepatitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hepatitis) - (1 / (a_hepatitis + b_hepatitis)) + (1 / c_hepatitis) - (1 / (c_hepatitis + d_hepatitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hepatitis / b_hepatitis) / (c_hepatitis / d_hepatitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hepatitis) + (1 / b_hepatitis) + (1 / c_hepatitis) + (1 / d_hepatitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hepatitis * (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis)) / ((a_hepatitis + c_hepatitis) * (a_hepatitis + b_hepatitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hepatitis - 1 / (a_hepatitis + b_hepatitis) - 1 / (a_hepatitis + c_hepatitis) + 1 / (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hepatitis)
  results_table_hepatitis <- rbind(results_table_hepatitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hepatitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hepatitis
print(results_table_hepatitis)

```
We repeat this with nephritis:
```{r}
# Filter for the specific drugs and count 'nephritis' reactions
nephritis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Nephritis = sum(c_across(starts_with("reaction_")) == "Nephritis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Nephritis_Count = sum(Nephritis))

# Print the result
print(nephritis_counts)
```
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

nephritis_counts <- nephritis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'nephritis' counts from total reactions
results_non_nephritis <- results_irae %>%
  mutate(
    Non_Nephritis_Reactions = Total_Reactions - nephritis_counts$Nephritis_Count
  )

# Print the updated results
print(results_non_nephritis)

# Create the new dataset by adding Non_Nephritis_Reactions to Total_Reactions
results_other_nephritis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_nephritis$Non_Nephritis_Reactions
)

# Print the new dataset
print(results_other_nephritis)
```
Disproportionality analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Nephritis
results_table_nephritis <- data.frame()

# Loop through each drug and perform the analysis for Nephritis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Nephritis)
  a_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Nephritis_Count)
  b_nephritis <- results_other_nephritis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Nephritis_Count)) %>% pull()
  d_nephritis <- results_other_nephritis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_nephritis / (a_nephritis + b_nephritis)) / (c_nephritis / (c_nephritis + d_nephritis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_nephritis) - (1 / (a_nephritis + b_nephritis)) + (1 / c_nephritis) - (1 / (c_nephritis + d_nephritis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_nephritis / b_nephritis) / (c_nephritis / d_nephritis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_nephritis) + (1 / b_nephritis) + (1 / c_nephritis) + (1 / d_nephritis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_nephritis * (a_nephritis + b_nephritis + c_nephritis + d_nephritis)) / ((a_nephritis + c_nephritis) * (a_nephritis + b_nephritis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_nephritis - 1 / (a_nephritis + b_nephritis) - 1 / (a_nephritis + c_nephritis) + 1 / (a_nephritis + b_nephritis + c_nephritis + d_nephritis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Nephritis)
  results_table_nephritis <- rbind(results_table_nephritis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Nephritis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Nephritis
print(results_table_nephritis)

```
```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(huxtable)

# Combine all individual results tables
all_results <- bind_rows(
  results_table_all_iraes,
  results_table_rash,
  results_table_pruritus,
  results_table_hypothyroidism,
  results_table_hyperthyroidism,
  results_table_colitis,
  results_table_hepatitis,
  results_table_nephritis
)

# Pivot data to get reactions as rows and drugs as grouped columns with subcolumns for PRR, ROR, and IC
summary_table <- all_results %>%
  pivot_wider(names_from = Drug, values_from = c(PRR, ROR, IC),
              names_glue = "{Drug} {.value} (95% CI)")

# Check the structure of the table and columns
colnames(summary_table)

# Highlight significant values using conditional bolding (PRR > 1, ROR > 1, IC > 0)
summary_table <- summary_table %>%
  mutate(across(starts_with("PRR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
  mutate(across(starts_with("ROR"), ~ ifelse(. > 1 & grepl("\\(1", .), "bold", .))) %>%
  mutate(across(starts_with("IC"), ~ ifelse(. > 0 & grepl("\\(0", .), "bold", .)))

# Split the summary_table into 4 separate tables for each drug
# Table for Atezolizumab
table_atezolizumab <- summary_table %>%
  select(Reaction, starts_with("Atezolizumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Atezolizumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Atezolizumab
for (i in 2:ncol(table_atezolizumab)) {
  table_atezolizumab <- table_atezolizumab %>%
    set_bold(which(table_atezolizumab[, i] == "bold"), i)
}

# Table for Avelumab
table_avelumab <- summary_table %>%
  select(Reaction, starts_with("Avelumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Avelumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Avelumab
for (i in 2:ncol(table_avelumab)) {
  table_avelumab <- table_avelumab %>%
    set_bold(which(table_avelumab[, i] == "bold"), i)
}

# Table for Nivolumab
table_nivolumab <- summary_table %>%
  select(Reaction, starts_with("Nivolumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Nivolumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Nivolumab
for (i in 2:ncol(table_nivolumab)) {
  table_nivolumab <- table_nivolumab %>%
    set_bold(which(table_nivolumab[, i] == "bold"), i)
}

# Table for Pembrolizumab
table_pembrolizumab <- summary_table %>%
  select(Reaction, starts_with("Pembrolizumab")) %>%
  as_hux() %>%
  set_bold(1, 1:ncol(.)) %>%
  set_caption("Pembrolizumab Drug-Reaction Analysis") %>%
  set_width(1)

# Apply bolding for significant values for Pembrolizumab
for (i in 2:ncol(table_pembrolizumab)) {
  table_pembrolizumab <- table_pembrolizumab %>%
    set_bold(which(table_pembrolizumab[, i] == "bold"), i)
}

# Display each table
table_atezolizumab
table_avelumab
table_nivolumab
table_pembrolizumab

```

```{r}
#Save the table
library(here)
# Load necessary libraries
library(here)

# Define paths for saving each table
table_atezolizumab_path <- here("results", "tables", "table_atezolizumab.rds")
table_avelumab_path <- here("results", "tables", "table_avelumab.rds")
table_nivolumab_path <- here("results", "tables", "table_nivolumab.rds")
table_pembrolizumab_path <- here("results", "tables", "table_pembrolizumab.rds")

# Save all four tables to their respective paths
saveRDS(table_atezolizumab, file = table_atezolizumab_path)
saveRDS(table_avelumab, file = table_avelumab_path)
saveRDS(table_nivolumab, file = table_nivolumab_path)
saveRDS(table_pembrolizumab, file = table_pembrolizumab_path)
```