### 2025-03-14 code/eda-code/eda.qmd

---
title: "Exploratory analysis"
---

We start of by summarizing all the information

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyr)
library(knitr)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

First, we summarize all our information using a table.

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(huxtable)

# 1. Summarize Age (Mean and SD)
age_summary <- mydata %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(
    mean_age = mean(patient_age, na.rm = TRUE),
    sd_age = sd(patient_age, na.rm = TRUE)
  ) %>%
  mutate(`Age, mean years ± SD` = paste0(round(mean_age, 1), " ± ", round(sd_age, 1))) %>%
  select(suspect_product_active_ingredients, `Age, mean years ± SD`)

# 2. Summarize Sex Distribution
sex_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, sex) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = sex, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Not Specified (Sex)` = `Not Specified`) %>%
  select(suspect_product_active_ingredients, Male, Female, `Not Specified (Sex)`)

# 3. Summarize Reaction Severity
reaction_severity_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, serious) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = serious, values_from = count, values_fill = list(count = 0)) %>%
  rename(Serious = Serious, `Non-Serious (Reaction)` = `Non-Serious`) %>%
  select(suspect_product_active_ingredients, Serious, `Non-Serious (Reaction)`)

# 4. Summarize Outcomes
outcomes_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, outcomes) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = outcomes, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Non-Serious (Outcome)` = `Non-Serious`) %>%
  select(suspect_product_active_ingredients, Died, Hospitalized, Disabled, `Life Threatening`, `Non-Serious (Outcome)`, `Other Outcomes`)

# 5. Summarize Reporter Type
reporter_type_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, reporter_type) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = reporter_type, values_from = count, values_fill = list(count = 0)) %>%
  select(suspect_product_active_ingredients, `Healthcare Professional`, Consumer)

# 6. Summarize Country
country_summary <- mydata %>%
  group_by(suspect_product_active_ingredients, country_where_event_occurred) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = country_where_event_occurred, values_from = count, values_fill = list(count = 0)) %>%
  rename(`Not Specified (Country)` = `Not Specified`) %>%
  select(suspect_product_active_ingredients, US, `Outside US`, `Not Specified (Country)`)

# Combine all summaries into one table
combined_summary <- age_summary %>%
  full_join(sex_summary, by = "suspect_product_active_ingredients") %>%
  full_join(reaction_severity_summary, by = "suspect_product_active_ingredients") %>%
  full_join(outcomes_summary, by = "suspect_product_active_ingredients") %>%
  full_join(reporter_type_summary, by = "suspect_product_active_ingredients") %>%
  full_join(country_summary, by = "suspect_product_active_ingredients")

# Convert all columns (except `suspect_product_active_ingredients`) to character
combined_summary <- combined_summary %>%
  mutate(across(-suspect_product_active_ingredients, as.character))

# Transpose the table for Word compatibility
transposed_table <- combined_summary %>%
  pivot_longer(cols = -suspect_product_active_ingredients, names_to = "Attribute", values_to = "Value") %>%
  pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value)

# Rename columns for clarity
colnames(transposed_table)[1] <- "Attribute"

# Convert to huxtable object
huxtable_table <- as_hux(transposed_table)

# Format the table with huxtable (style, borders, etc.)
huxtable_table <- huxtable_table %>%
  set_all_borders(1) %>%
  set_font_size(12) %>%
  set_align("center") %>%
  set_width(1.5)

# Print the huxtable object, this will render correctly in Word
huxtable_table

```

We save the table.

```{r}
# save to file
huxtable_table_path = here("results", "tables", "summarytable.rds")
saveRDS(huxtable_table, file = huxtable_table_path)
```

Table 2 - Summarizing organ system reaction by drug type
```{r}
# Load necessary libraries
library(huxtable)
library(dplyr)
library(tidyr)

# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

# Transform the organ system columns into a single column and calculate counts and percentages
organ_system_summary <- mydata %>%
    select(suspect_product_active_ingredients, all_of(organ_systems)) %>%
    pivot_longer(cols = all_of(organ_systems), names_to = "Organ_System", values_to = "Reaction") %>%
    filter(!is.na(Reaction)) %>%
    group_by(suspect_product_active_ingredients, Reaction) %>%
    summarise(Count = n(), .groups = 'drop') %>%
    mutate(Percent = round((Count / nrow(mydata)) * 100, 1)) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# Pivot the summary to have reactions as rows and drugs as columns, combining counts and percentages
organ_system_pivoted <- organ_system_summary %>%
    select(suspect_product_active_ingredients, Reaction, Value) %>%
    pivot_wider(names_from = suspect_product_active_ingredients, values_from = Value, values_fill = list(Value = "0"))

# Create a huxtable table
ht <- huxtable::as_hux(organ_system_pivoted)

# Add styling to the table
ht <- ht %>%
    set_bold(1, 1:ncol(ht)) %>%  # Bold the first row (header)
    set_caption("Immune Checkpoint Inhibitors by Organ System and Reaction") %>%
    set_width(1)                  # Set column width to make the table more compact

# Display the table
ht

#Save the table
ht_path = here("results", "tables", "summarytable1.rds")
saveRDS(ht, file = ht_path)

```

# Data exploration through figures

Bar graph for outcomes by drugs.

```{r}
library(ggplot2)

# Define custom colors for the 'serious' variable
custom_colors <- c(
    "Serious" = "#d7191c",  # Red for serious
    "Non-Serious" = "#2c7bb6"    # Blue for non-serious
)

# Create the plot with custom colors for 'serious'
p1 <- mydata %>%
    ggplot(aes(x = suspect_product_active_ingredients, fill = serious)) +
    geom_bar(position = "stack") + # Use "stack" for stacked bars
    scale_fill_manual(values = custom_colors) + # Apply custom colors
    labs(
        title = "Bar Plot of Outcomes of Immune Checkpoint Inhibitors",
        x = "Immune Checkpoint Inhibitors",
        y = "Count",
        fill = "Outcomes"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

# Print the plot
print(p1)

figure_file = here("results", "figures","figure1.png")
ggsave(filename = figure_file, plot=p1) 
```

Now let's look at the bar graph by type of organ system.

```{r}
# Define the organ system columns
organ_systems <- c("organ_system_1", "organ_system_2", "organ_system_3", "organ_system_4", "organ_system_5")

# Create a summary dataset for reactions across all organ systems
organ_system_summary <- mydata %>%
    select(all_of(organ_systems)) %>%
    pivot_longer(cols = everything(), names_to = "Organ_System", values_to = "Reaction") %>%
    filter(!is.na(Reaction)) %>%
    group_by(Reaction) %>%
    summarise(Count = n(), .groups = 'drop') %>%
    mutate(Percent = round((Count / sum(Count)) * 100, 1)) %>%
    mutate(Value = paste(Count, " (", Percent, "%)", sep = ""))

# View the summary dataset
organ_system_summary

# Plot the reaction types with counts and percentages
p2 <- organ_system_summary %>%
    ggplot(aes(x = reorder(Reaction, Count), y = Count)) +  # Reorder by Count
    geom_col(fill = "#1f77b4") +
    labs(
        title = "Bar Plot of Reaction Types",
        x = "Reaction Type",
        y = "Count",
        caption = "Percentages are shown in parentheses"
    ) +
    theme_minimal() +
    geom_text(aes(label = Value), hjust = 1.1, size = 3) +  # Adjust label position to avoid overlap
    coord_flip() +  # Flip the axes to make it horizontal
    theme(
        plot.margin = margin(10, 50, 10, 10)  # Add margin to the right side to avoid label cut-off
    )

# Print the plot
print(p2)

figure_file = here("results", "figures","figure2.png")
ggsave(filename = figure_file, plot=p2) 
```