### 2025-04-10 code/analysis-code/statistical-analysis.qmd
Updates to Part 4 of MADA Project

---
title: "statistical-analysis"
format: html
---

This file contains details of the statistical analysis.

After the exploratory data analysis and the descriptive statistics, the next step is to run a disproportionality analysis. The first step is to assess the adverse events in comparison to the entire FAERS database.

To assess this, we need the following information: 
a = Number of irAEs with 1 ICI (e.g. atezolizumab) 
b = Number of reactions other than irAEs with atezolizumab
c = Number of irAEs with drugs other than atezolizumab 
d = Number of reactions other than irAEs with drugs other than atzeolizumab.

The same procedure would be repeated for the other 3 ICIs.

First, we load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

First we find number of irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions <- function(data, drug) {
  # Filter rows for the drug
  drug_data <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions <- c(
    drug_data$reaction_1,
    drug_data$reaction_2,
    drug_data$reaction_3,
    drug_data$reaction_4,
    drug_data$reaction_5
  )

  # Remove NA values
  all_reactions <- all_reactions[!is.na(all_reactions)]

  # Count total reactions
  total_reactions <- length(all_reactions)

  # Return total reactions
  return(total_reactions)
}

# Count total reactions for each drug
results <- sapply(drugs, function(drug) count_total_reactions(mydata, drug))

# Print the results
results_irae <- data.frame(Drug = drugs, Total_Reactions = results)
print(results_irae)
```

The `echo: false` option disables the printing of code (only output is displayed).

Now, we do the same for reactions other than irAEs. We created and saved this data when we processed it.

Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata1.rds")
#load data
mydata1 <- readRDS(data_location)
```

Now we find number of reactions other than irAEs with each ICI.

```{r}
# Load necessary library
library(dplyr)

# List of drugs to analyze
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Function to count total reactions for a given drug
count_total_reactions1 <- function(data, drug) {
  # Filter rows for the drug
  drug_data1 <- data %>%
    filter(grepl(drug, suspect_product_active_ingredients, ignore.case = TRUE))

  # Combine reaction columns into a single vector
  all_reactions1 <- c(
    drug_data1$reaction_1,
    drug_data1$reaction_2,
    drug_data1$reaction_3,
    drug_data1$reaction_4,
    drug_data1$reaction_5
  )

  # Remove NA values
  all_reactions1 <- all_reactions1[!is.na(all_reactions1)]

  # Count total reactions
  total_reactions1 <- length(all_reactions1)

  # Return total reactions
  return(total_reactions1)
}

# Count total reactions for each drug
results1 <- sapply(drugs, function(drug) count_total_reactions1(mydata1, drug))

# Print the results
results_non_irae <- data.frame(Drug = drugs, Total_Reactions = results1)
print(results_non_irae)
```
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for All irAEs
results_table_all_iraes <- data.frame()

# Loop through each drug and perform the analysis for all irAEs
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for all irAEs combined)
  a <- results_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  b <- results_non_irae %>% filter(Drug == drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  c <- results_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()
  d <- results_non_irae %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a / (a + b)) / (c / (c + d))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a) - (1 / (a + b)) + (1 / c) - (1 / (c + d)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a / b) / (c / d)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a) + (1 / b) + (1 / c) + (1 / d))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a * (a + b + c + d)) / ((a + c) * (a + b)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a - 1 / (a + b) - 1 / (a + c) + 1 / (a + b + c + d))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (All irAEs)
  results_table_all_iraes <- rbind(results_table_all_iraes, data.frame(
    Drug = drug_of_interest,
    Reaction = "All irAEs",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for All irAEs
print(results_table_all_iraes)
```
The next step is to assess the safety signal for individual irAEs.
The commonly reported irAEs in clinical trials are rash, pruritus, colitis, hepatitis, hyperthyroidism, hypothyroidism, and nephritis.
We conduct disproportionality analysis for each of these.

We compare each of these reactions to all other reactions.
First, we assess the number of 'rash' reactions reported with all the ICIs.

```{r}
 # Filter for the specific drugs and count 'Rash' reactions
rash_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Rash = sum(c_across(starts_with("reaction_")) == "Rash", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Rash_Count = sum(Rash))

# Print the result
print(rash_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'rash' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

rash_counts <- rash_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Rash' counts from total reactions
results_non_rash <- results_irae %>%
  mutate(
    Non_Rash_Reactions = Total_Reactions - rash_counts$Rash_Count
  )

# Print the updated results
print(results_non_rash)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_rash <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_rash$Non_Rash_Reactions
)

# Print the new dataset
print(results_other_rash)

```

We have all our numbers now.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Rash
results_table_rash <- data.frame()

# Loop through each drug and perform the analysis for Rash
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Rash)
  a_rash <- rash_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Rash_Count)
  b_rash <- results_other_rash %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_rash <- rash_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Rash_Count)) %>% pull()
  d_rash <- results_other_rash %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_rash / (a_rash + b_rash)) / (c_rash / (c_rash + d_rash))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_rash) - (1 / (a_rash + b_rash)) + (1 / c_rash) - (1 / (c_rash + d_rash)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_rash / b_rash) / (c_rash / d_rash)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_rash) + (1 / b_rash) + (1 / c_rash) + (1 / d_rash))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_rash * (a_rash + b_rash + c_rash + d_rash)) / ((a_rash + c_rash) * (a_rash + b_rash)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_rash - 1 / (a_rash + b_rash) - 1 / (a_rash + c_rash) + 1 / (a_rash + b_rash + c_rash + d_rash))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Rash)
  results_table_rash <- rbind(results_table_rash, data.frame(
    Drug = drug_of_interest,
    Reaction = "Rash",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Rash
print(results_table_rash)

```

Now, we conduct the same analysis for pruritus.
```{r}
 # Filter for the specific drugs and count 'Pruritus' reactions
pruritus_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Pruritus = sum(c_across(starts_with("reaction_")) == "Pruritus", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Pruritus_Count = sum(Pruritus))

# Print the result
print(pruritus_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'pruritus' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

pruritus_counts <- pruritus_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'pruritus' counts from total reactions
results_non_pruritus <- results_irae %>%
  mutate(
    Non_Pruritus_Reactions = Total_Reactions - pruritus_counts$Pruritus_Count
  )

# Print the updated results
print(results_non_pruritus)

# Create the new dataset by adding Non_Pruritus_Reactions to Total_Reactions
results_other_pruritus <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_pruritus$Non_Pruritus_Reactions
)

# Print the new dataset
print(results_other_pruritus)

```
We now conduct the analysis.

```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Pruritus
results_table_pruritus <- data.frame()

# Loop through each drug and perform the analysis for Pruritus
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Pruritus)
  a_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Pruritus_Count)
  b_pruritus <- results_other_pruritus %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_pruritus <- pruritus_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Pruritus_Count)) %>% pull()
  d_pruritus <- results_other_pruritus %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_pruritus / (a_pruritus + b_pruritus)) / (c_pruritus / (c_pruritus + d_pruritus))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_pruritus) - (1 / (a_pruritus + b_pruritus)) + (1 / c_pruritus) - (1 / (c_pruritus + d_pruritus)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_pruritus / b_pruritus) / (c_pruritus / d_pruritus)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_pruritus) + (1 / b_pruritus) + (1 / c_pruritus) + (1 / d_pruritus))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_pruritus * (a_pruritus + b_pruritus + c_pruritus + d_pruritus)) / ((a_pruritus + c_pruritus) * (a_pruritus + b_pruritus)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_pruritus - 1 / (a_pruritus + b_pruritus) - 1 / (a_pruritus + c_pruritus) + 1 / (a_pruritus + b_pruritus + c_pruritus + d_pruritus))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Pruritus)
  results_table_pruritus <- rbind(results_table_pruritus, data.frame(
    Drug = drug_of_interest,
    Reaction = "Pruritus",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Pruritus
print(results_table_pruritus)

```

We repeat this analysis for hypothyroidism.
```{r}
 # Filter for the specific drugs and count 'Hypothyroidism' reactions
hypothyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hypothyroidism = sum(c_across(starts_with("reaction_")) == "Hypothyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hypothyroidism_Count = sum(Hypothyroidism))

# Print the result
print(hypothyroidism_counts)
```

We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hypothyroidism' reactions associated with each ICI.

```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hypothyroidism_counts <- hypothyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hypothyroidism' counts from total reactions
results_non_hypothyroidism <- results_irae %>%
  mutate(
    Non_Hypothyroidism_Reactions = Total_Reactions - hypothyroidism_counts$Hypothyroidism_Count
  )

# Print the updated results
print(results_non_hypothyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hypothyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hypothyroidism$Non_Hypothyroidism_Reactions
)

# Print the new dataset
print(results_other_hypothyroidism)

```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hypothyroidism
results_table_hypothyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hypothyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hypothyroidism)
  a_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hypothyroidism_Count)
  b_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hypothyroidism <- hypothyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hypothyroidism_Count)) %>% pull()
  d_hypothyroidism <- results_other_hypothyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hypothyroidism / (a_hypothyroidism + b_hypothyroidism)) / (c_hypothyroidism / (c_hypothyroidism + d_hypothyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hypothyroidism) - (1 / (a_hypothyroidism + b_hypothyroidism)) + (1 / c_hypothyroidism) - (1 / (c_hypothyroidism + d_hypothyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hypothyroidism / b_hypothyroidism) / (c_hypothyroidism / d_hypothyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hypothyroidism) + (1 / b_hypothyroidism) + (1 / c_hypothyroidism) + (1 / d_hypothyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hypothyroidism * (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism)) / ((a_hypothyroidism + c_hypothyroidism) * (a_hypothyroidism + b_hypothyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hypothyroidism - 1 / (a_hypothyroidism + b_hypothyroidism) - 1 / (a_hypothyroidism + c_hypothyroidism) + 1 / (a_hypothyroidism + b_hypothyroidism + c_hypothyroidism + d_hypothyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hypothyroidism)
  results_table_hypothyroidism <- rbind(results_table_hypothyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hypothyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hypothyroidism
print(results_table_hypothyroidism)

```

Now, we assess the reports for hyperthyroidism:
```{r}
 # Filter for the specific drugs and count 'hyperthyroidism' reactions
hyperthyroidism_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Hyperthyroidism = sum(c_across(starts_with("reaction_")) == "Hyperthyroidism", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Hyperthyroidism_Count = sum(Hyperthyroidism))

# Print the result
print(hyperthyroidism_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hyperthyroidism' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hyperthyroidism_counts <- hyperthyroidism_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'Hyperthyroidism' counts from total reactions
results_non_hyperthyroidism <- results_irae %>%
  mutate(
    Non_Hyperthyroidism_Reactions = Total_Reactions - hyperthyroidism_counts$Hyperthyroidism_Count
  )

# Print the updated results
print(results_non_hyperthyroidism)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_hyperthyroidism <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hyperthyroidism$Non_Hyperthyroidism_Reactions
)

# Print the new dataset
print(results_other_hyperthyroidism)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hyperthyroidism
results_table_hyperthyroidism <- data.frame()

# Loop through each drug and perform the analysis for Hyperthyroidism
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hyperthyroidism)
  a_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hyperthyroidism_Count)
  b_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hyperthyroidism <- hyperthyroidism_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hyperthyroidism_Count)) %>% pull()
  d_hyperthyroidism <- results_other_hyperthyroidism %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hyperthyroidism / (a_hyperthyroidism + b_hyperthyroidism)) / (c_hyperthyroidism / (c_hyperthyroidism + d_hyperthyroidism))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hyperthyroidism) - (1 / (a_hyperthyroidism + b_hyperthyroidism)) + (1 / c_hyperthyroidism) - (1 / (c_hyperthyroidism + d_hyperthyroidism)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hyperthyroidism / b_hyperthyroidism) / (c_hyperthyroidism / d_hyperthyroidism)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hyperthyroidism) + (1 / b_hyperthyroidism) + (1 / c_hyperthyroidism) + (1 / d_hyperthyroidism))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hyperthyroidism * (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism)) / ((a_hyperthyroidism + c_hyperthyroidism) * (a_hyperthyroidism + b_hyperthyroidism)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hyperthyroidism - 1 / (a_hyperthyroidism + b_hyperthyroidism) - 1 / (a_hyperthyroidism + c_hyperthyroidism) + 1 / (a_hyperthyroidism + b_hyperthyroidism + c_hyperthyroidism + d_hyperthyroidism))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hyperthyroidism)
  results_table_hyperthyroidism <- rbind(results_table_hyperthyroidism, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hyperthyroidism",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hyperthyroidism
print(results_table_hyperthyroidism)

```

Now, we repeat this analysis with colitis:

```{r}
 # Filter for the specific drugs and count 'colitis' reactions
colitis_counts <- mydata %>%
     filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
     rowwise() %>%
     mutate(Colitis = sum(c_across(starts_with("reaction_")) == "Colitis", na.rm = TRUE)) %>%
     group_by(suspect_product_active_ingredients) %>%
     summarise(Colitis_Count = sum(Colitis))

# Print the result
print(colitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'colitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

colitis_counts <- colitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'colitis' counts from total reactions
results_non_colitis <- results_irae %>%
  mutate(
    Non_Colitis_Reactions = Total_Reactions - colitis_counts$Colitis_Count
  )

# Print the updated results
print(results_non_colitis)

# Create the new dataset by adding Non_Rash_Reactions to Total_Reactions
results_other_colitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_colitis$Non_Colitis_Reactions
)

# Print the new dataset
print(results_other_colitis)
```

Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Colitis
results_table_colitis <- data.frame()

# Loop through each drug and perform the analysis for Colitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Colitis)
  a_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Colitis_Count)
  b_colitis <- results_other_colitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_colitis <- colitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Colitis_Count)) %>% pull()
  d_colitis <- results_other_colitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_colitis / (a_colitis + b_colitis)) / (c_colitis / (c_colitis + d_colitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_colitis) - (1 / (a_colitis + b_colitis)) + (1 / c_colitis) - (1 / (c_colitis + d_colitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_colitis / b_colitis) / (c_colitis / d_colitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_colitis) + (1 / b_colitis) + (1 / c_colitis) + (1 / d_colitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_colitis * (a_colitis + b_colitis + c_colitis + d_colitis)) / ((a_colitis + c_colitis) * (a_colitis + b_colitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_colitis - 1 / (a_colitis + b_colitis) - 1 / (a_colitis + c_colitis) + 1 / (a_colitis + b_colitis + c_colitis + d_colitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Colitis)
  results_table_colitis <- rbind(results_table_colitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Colitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Colitis
print(results_table_colitis)

```
Now, we conduct the same analysis for hepatitis:

```{r}
# Filter for the specific drugs and count 'hepatitis' reactions
hepatitis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Hepatitis = sum(c_across(starts_with("reaction_")) == "Hepatitis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Hepatitis_Count = sum(Hepatitis))

# Print the result
print(hepatitis_counts)
```
We subtract this number from the irAEs, and add this number to the non-irAEs so as to get the total number of non-'hepatitis' reactions associated with each ICI.
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

hepatitis_counts <- hepatitis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'hepatitis' counts from total reactions
results_non_hepatitis <- results_irae %>%
  mutate(
    Non_Hepatitis_Reactions = Total_Reactions - hepatitis_counts$Hepatitis_Count
  )

# Print the updated results
print(results_non_hepatitis)

# Create the new dataset by adding Non_Hepatitis_Reactions to Total_Reactions
results_other_hepatitis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_hepatitis$Non_Hepatitis_Reactions
)

# Print the new dataset
print(results_other_hepatitis)
```
Now, we conduct the analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Hepatitis
results_table_hepatitis <- data.frame()

# Loop through each drug and perform the analysis for Hepatitis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Hepatitis)
  a_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Hepatitis_Count)
  b_hepatitis <- results_other_hepatitis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_hepatitis <- hepatitis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Hepatitis_Count)) %>% pull()
  d_hepatitis <- results_other_hepatitis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_hepatitis / (a_hepatitis + b_hepatitis)) / (c_hepatitis / (c_hepatitis + d_hepatitis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_hepatitis) - (1 / (a_hepatitis + b_hepatitis)) + (1 / c_hepatitis) - (1 / (c_hepatitis + d_hepatitis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_hepatitis / b_hepatitis) / (c_hepatitis / d_hepatitis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_hepatitis) + (1 / b_hepatitis) + (1 / c_hepatitis) + (1 / d_hepatitis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_hepatitis * (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis)) / ((a_hepatitis + c_hepatitis) * (a_hepatitis + b_hepatitis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_hepatitis - 1 / (a_hepatitis + b_hepatitis) - 1 / (a_hepatitis + c_hepatitis) + 1 / (a_hepatitis + b_hepatitis + c_hepatitis + d_hepatitis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Hepatitis)
  results_table_hepatitis <- rbind(results_table_hepatitis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Hepatitis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Hepatitis
print(results_table_hepatitis)

```
We repeat this with nephritis:
```{r}
# Filter for the specific drugs and count 'nephritis' reactions
nephritis_counts <- mydata %>%
  filter(suspect_product_active_ingredients %in% c("Atezolizumab", "Avelumab", "Pembrolizumab", "Nivolumab")) %>%
  rowwise() %>%
  mutate(Nephritis = sum(c_across(starts_with("reaction_")) == "Nephritis", na.rm = TRUE)) %>%
  group_by(suspect_product_active_ingredients) %>%
  summarise(Nephritis_Count = sum(Nephritis))

# Print the result
print(nephritis_counts)
```
```{r}
# Ensure the data frames are in the same order
results_irae <- results_irae %>%
  arrange(Drug)

nephritis_counts <- nephritis_counts %>%
  arrange(suspect_product_active_ingredients)

# Subtract 'nephritis' counts from total reactions
results_non_nephritis <- results_irae %>%
  mutate(
    Non_Nephritis_Reactions = Total_Reactions - nephritis_counts$Nephritis_Count
  )

# Print the updated results
print(results_non_nephritis)

# Create the new dataset by adding Non_Nephritis_Reactions to Total_Reactions
results_other_nephritis <- data.frame(
  Drug = results_non_irae$Drug,
  Total_Reactions = results_non_irae$Total_Reactions + results_non_nephritis$Non_Nephritis_Reactions
)

# Print the new dataset
print(results_other_nephritis)
```
Disproportionality analysis:
```{r}
# Load necessary library
library(dplyr)

# Define the list of drugs
drugs <- c("Atezolizumab", "Avelumab", "Nivolumab", "Pembrolizumab")

# Initialize an empty data frame to store results for Nephritis
results_table_nephritis <- data.frame()

# Loop through each drug and perform the analysis for Nephritis
for (drug_of_interest in drugs) {

  # Extract a, b, c, d for the current drug (for Nephritis)
  a_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients == drug_of_interest) %>% pull(Nephritis_Count)
  b_nephritis <- results_other_nephritis %>% filter(Drug == drug_of_interest) %>% pull(Total_Reactions)
  c_nephritis <- nephritis_counts %>% filter(suspect_product_active_ingredients != drug_of_interest) %>% summarise(sum(Nephritis_Count)) %>% pull()
  d_nephritis <- results_other_nephritis %>% filter(Drug != drug_of_interest) %>% summarise(sum(Total_Reactions)) %>% pull()

  # Step 1: Calculate PRR (Proportional Reporting Ratio)
  prr <- (a_nephritis / (a_nephritis + b_nephritis)) / (c_nephritis / (c_nephritis + d_nephritis))

  # Step 2: Calculate the 95% confidence interval for PRR
  se_log_prr <- sqrt((1 / a_nephritis) - (1 / (a_nephritis + b_nephritis)) + (1 / c_nephritis) - (1 / (c_nephritis + d_nephritis)))
  prr_lower <- exp(log(prr) - 1.96 * se_log_prr)
  prr_upper <- exp(log(prr) + 1.96 * se_log_prr)

  # Step 3: Calculate ROR (Reporting Odds Ratio)
  ror <- (a_nephritis / b_nephritis) / (c_nephritis / d_nephritis)

  # Step 4: Calculate the 95% confidence interval for ROR
  se_log_ror <- sqrt((1 / a_nephritis) + (1 / b_nephritis) + (1 / c_nephritis) + (1 / d_nephritis))
  ror_lower <- exp(log(ror) - 1.96 * se_log_ror)
  ror_upper <- exp(log(ror) + 1.96 * se_log_ror)

  # Step 5: Calculate IC (Information Component)
  IC <- log2((a_nephritis * (a_nephritis + b_nephritis + c_nephritis + d_nephritis)) / ((a_nephritis + c_nephritis) * (a_nephritis + b_nephritis)))

  # Step 6: Calculate the variance of IC
  var_IC <- (1 / log(2)^2) * (1 / a_nephritis - 1 / (a_nephritis + b_nephritis) - 1 / (a_nephritis + c_nephritis) + 1 / (a_nephritis + b_nephritis + c_nephritis + d_nephritis))

  # Step 7: Calculate the 95% confidence interval for IC
  IC_lower <- IC - 1.96 * sqrt(var_IC)
  IC_upper <- IC + 1.96 * sqrt(var_IC)

  # Store results for the current drug (Nephritis)
  results_table_nephritis <- rbind(results_table_nephritis, data.frame(
    Drug = drug_of_interest,
    Reaction = "Nephritis",
    PRR = paste0(round(prr, 2), " (", round(prr_lower, 2), ", ", round(prr_upper, 2), ")"),
    ROR = paste0(round(ror, 2), " (", round(ror_lower, 2), ", ", round(ror_upper, 2), ")"),
    IC = paste0(round(IC, 2), " (", round(IC_lower, 2), ", ", round(IC_upper, 2), ")")
  ))
}

# Print the results table for Nephritis
print(results_table_nephritis)

```
```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(huxtable)

# Combine all individual results tables
all_results <- bind_rows(
  results_table_all_iraes,
  results_table_rash,
  results_table_pruritus,
  results_table_hypothyroidism,
  results_table_hyperthyroidism,
  results_table_colitis,
  results_table_hepatitis,
  results_table_nephritis
)

# Pivot data to get reactions as rows and drugs as grouped columns
summary_table <- all_results %>%
  pivot_wider(
    names_from = Drug, 
    values_from = c(PRR, ROR, IC),
    names_glue = "{Drug} {.value} (95% CI)"
  )

# Function to create formatted tables with WORKING conditional formatting
create_drug_table <- function(drug_name, summary_data) {
  # Select relevant columns
  drug_table <- summary_data %>%
    select(Reaction, starts_with(drug_name)) %>%
    as_hux(add_colnames = TRUE)

  # Basic formatting
  drug_table <- drug_table %>%
    set_bold(1, everywhere, TRUE) %>%          # Bold headers
    set_align(everywhere, everywhere, "center") %>%
    set_all_borders(brdr(1, "solid")) %>%      # Solid borders
    set_all_padding(4) %>%                     # Cell padding
    set_font_size(10) %>%                      # Font size
    set_width(0.9)                             # Table width

  # Add caption as merged header row
  drug_table <- insert_row(drug_table, 
                         c(paste(drug_name, "Drug-Reaction Analysis"), 
                           rep("", ncol(drug_table)-1)), 
                         after = 0) %>%
    set_colspan(1, 1, ncol(drug_table)) %>%
    set_bold(1, everywhere, TRUE) %>%          # Bold title
    set_background_color(1, everywhere, "gray95") %>%
    set_font_size(1, everywhere, 12)

  # Apply conditional formatting - SIMPLIFIED AND WORKING VERSION
  if (ncol(drug_table) > 1) {
    for (i in 2:ncol(drug_table)) {
      # Get the column type (PRR, ROR, or IC)
      col_type <- colnames(drug_table)[i]

      # Process each row (skip header row)
      for (j in 2:nrow(drug_table)) {
        cell_value <- as.character(drug_table[j, i])

        # Extract the first number (point estimate)
        point_est <- as.numeric(gsub("([0-9.]+).*", "\\1", cell_value))

        # Extract the first number in parentheses (CI lower bound)
        ci_lower <- as.numeric(gsub(".*\\(([0-9.]+).*", "\\1", cell_value))

        # Check significance
        if (!is.na(point_est) && !is.na(ci_lower)) {
          if (grepl("PRR|ROR", col_type) && point_est > 1 && ci_lower > 1) {
            drug_table <- set_bold(drug_table, j, i, TRUE)
          } else if (grepl("IC", col_type) && point_est > 0 && ci_lower > 0) {
            drug_table <- set_bold(drug_table, j, i, TRUE)
          }
        }
      }
    }
  }

  return(drug_table)
}

# Create all tables
table_atezolizumab <- create_drug_table("Atezolizumab", summary_table)
table_avelumab <- create_drug_table("Avelumab", summary_table)
table_nivolumab <- create_drug_table("Nivolumab", summary_table)
table_pembrolizumab <- create_drug_table("Pembrolizumab", summary_table)

# Display tables
table_atezolizumab
table_avelumab
table_nivolumab
table_pembrolizumab 
```

```{r}
#Save the table
library(here)
# Load necessary libraries
library(here)

# Define paths for saving each table
table_atezolizumab_path <- here("results", "tables", "table_atezolizumab.rds")
table_avelumab_path <- here("results", "tables", "table_avelumab.rds")
table_nivolumab_path <- here("results", "tables", "table_nivolumab.rds")
table_pembrolizumab_path <- here("results", "tables", "table_pembrolizumab.rds")

# Save all four tables to their respective paths
saveRDS(table_atezolizumab, file = table_atezolizumab_path)
saveRDS(table_avelumab, file = table_avelumab_path)
saveRDS(table_nivolumab, file = table_nivolumab_path)
saveRDS(table_pembrolizumab, file = table_pembrolizumab_path)
```

First, let's fit a logistic regression model to see which factors affect the severity of the reactions:

Let's look at the relationship between severity and sex. We need to recode the sex variable for this:
```{r}
# Create a new binary column: 1 = Serious, 0 = Non-Serious
mydata$serious_bin <- ifelse(mydata$serious == "Serious", 1, 0)
```

Now, we fit the model and look at the plot.

```{r}
library(ggplot2)

logit_model_sex <- glm(serious_bin ~ sex, data = mydata, family = binomial)
summary(logit_model_sex)
mydata$predicted_prob <- predict(logit_model_sex, type = "response")

severity_sex = ggplot(mydata, aes(x = sex, y = predicted_prob)) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "purple") +
  labs(
    title = "Predicted Probability of Serious Event by Sex",
    x = "Sex",
    y = "Predicted Probability"
  ) +
  theme_minimal()

print(severity_sex)

# Define paths
severity_sex_path <- here("results", "figures", "severity_sex.png")
logit_model_sex_path = here("results", "tables", "logit_model_sex.rds")

# Save all four tables to their respective paths
ggsave(filename = severity_sex_path, plot=severity_sex) 
saveRDS(logit_model_sex, file = logit_model_sex_path)

```

We see that the probability of serious outcomes is highest for sex not specified, followed by males, then females.
There are more cases for males than females, so it is expected that more males report severe cases.
However, it is surprising that even with a low percentage of non-specified sex reports, the severity can be high.
This may indicate underlying issues with the data reporting.
We do see that sex is not a significant predictor of severity.

Let's look at the association of age with severity.
We know that age has a few NA's. Since, this is continuous data, it is better to remove the NA's.

```{r}
mydata_clean <- mydata[!is.na(mydata$patient_age), ]
```

```{r}
logit_model_age <- glm(serious_bin ~ patient_age, data = mydata_clean, family = binomial)
summary(logit_model_age)
mydata_clean$predicted_prob <- predict(logit_model_age, type = "response")

severity_age = ggplot(mydata_clean, aes(x = patient_age, y = predicted_prob)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  labs(
    title = "Predicted Probability of Serious Outcome by Age",
    x = "Patient Age",
    y = "Predicted Probability"
  ) +
  theme_minimal()

print(severity_age)

# Define paths
severity_age_path <- here("results", "figures", "severity_age.png")
logit_model_age_path = here("results", "tables", "logit_model_age.rds")

# Save all four tables to their respective paths
ggsave(filename = severity_age_path, plot=severity_age) 
saveRDS(logit_model_age, file = logit_model_age_path)

```
We see that predicted severity decreases with age.
However, this relationship is also not signficant.

Now, we use machine learning methods (random forest) to assess severity of reaction based on age, sex and drug.
We know that there are NA's in the age.
Let's impute the data (since removing the variables will create a very small dataset).
There were too many NA's for KNN imputation to be used.

```{r}
# Calculate the mean of the 'patient_age' column, excluding missing values
mean_age <- mean(mydata$patient_age, na.rm = TRUE)

# Replace missing values (NA) in 'patient_age' with the calculated mean
mydata$patient_age_imputed <- ifelse(is.na(mydata$patient_age), mean_age, mydata$patient_age)

```

First, let's split the data:

```{r}
set.seed(123)
library(caret)

train_index <- createDataPartition(mydata$serious_bin, p = 0.7, list = FALSE)
train_data <- mydata[train_index, ]
test_data  <- mydata[-train_index, ]

```

First, let's ensure that the serious variable is a factor:
```{r}
# Ensure 'serious_bin' and 'sex' is a factor
train_data$serious_bin <- as.factor(train_data$serious_bin)
test_data$serious_bin <- as.factor(test_data$serious_bin)

train_data$sex <- factor(train_data$sex)
test_data$sex <- factor(test_data$sex)
```

First, let's use a basic logistic regression model:
```{r}
# Fit logistic regression model with cross-validation
logit_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
                     data = train_data, 
                     method = "glm", 
                     family = "binomial", 
                     trControl = trainControl(method = "cv", number = 5))

```

Now, let's fit a decision tree (for classification):
```{r}
# Decision Tree (CART) model with cross-validation
dt_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
                  data = train_data, 
                  method = "rpart", 
                  trControl = trainControl(method = "cv", number = 5))

```

Now, let's use a random forest:
```{r}
# Random Forest model with cross-validation
rf_model <- train(serious_bin ~ patient_age_imputed + sex + suspect_product_active_ingredients, 
                  data = train_data, 
                  method = "rf", 
                  trControl = trainControl(method = "cv", number = 5))

```

Now, let's evaluate all the models:

```{r}
# Accuracy and AUC for Logistic Regression
logit_model$results

# Accuracy and AUC for Decision Tree
dt_model$results

# Accuracy and AUC for Random Forest
rf_model$results

```

```{r}
library(pROC)
library(ggplot2)
library(here)

# Logistic Regression ROC
logit_roc <- roc(train_data$serious_bin, predict(logit_model, type = "prob")[,2])
logit_roc_plot <- ggroc(logit_roc) + ggtitle("Logistic Regression ROC Curve")
cat("Logistic Regression AUC:", auc(logit_roc), "\n")

# Decision Tree ROC
dt_roc <- roc(train_data$serious_bin, predict(dt_model, type = "prob")[,2])
dt_roc_plot <- ggroc(dt_roc) + ggtitle("Decision Tree ROC Curve")
cat("Decision Tree AUC:", auc(dt_roc), "\n")

# Random Forest ROC
rf_roc <- roc(train_data$serious_bin, predict(rf_model, type = "prob")[,2])
rf_roc_plot <- ggroc(rf_roc) + ggtitle("Random Forest ROC Curve")
cat("Random Forest AUC:", auc(rf_roc), "\n")

# Define file paths
logit_roc_path <- here("results", "figures", "logit_roc.png")
dt_roc_path <- here("results", "figures", "dt_roc.png")
rf_roc_path <- here("results", "figures", "rf_roc.png")

# Save plots
ggsave(filename = logit_roc_path, plot = logit_roc_plot, width = 6, height = 4)
ggsave(filename = dt_roc_path, plot = dt_roc_plot, width = 6, height = 4)
ggsave(filename = rf_roc_path, plot = rf_roc_plot, width = 6, height = 4)

```

```{r}
cat("\nModel Comparison based on AUC:\n")
cat("Logistic Regression AUC:", auc(logit_roc), "\n")
cat("Decision Tree AUC:", auc(dt_roc), "\n")
cat("Random Forest AUC:", auc(rf_roc), "\n")
```

We see that the Random Forest model has an AUC of 0.806. Thus, this models is able to predict between serious and non-serious cases ~81% of the time.
The logistic regression model is next, able to distinguish ~60% of the time.
The decision tree model performs poorly, possibly due to fitting issues.

Let's fit the random forest model to the test data:
```{r}
# Make predictions on the test data (probabilities for AUC)
rf_pred_prob <- predict(rf_model, newdata = test_data, type = "prob")

# Make predictions on the test data (class labels for evaluation)
rf_pred_class <- predict(rf_model, newdata = test_data, type = "raw")

# ROC curve for Random Forest
rf_roc <- roc(test_data$serious_bin, rf_pred_prob[, 2])  # Using the probability of the 'Serious' class
plot(rf_roc, main = "Random Forest ROC Curve")
cat("Random Forest AUC on Test Data:", auc(rf_roc), "\n")

# Extract the random forest model from the train object
rf_final_model <- rf_model$finalModel

library(randomForest)
# Display feature importance
importance(rf_final_model)

# Plot feature importance
varImpPlot(rf_final_model)

```

Save the file:
```{r}
# Load library
library(randomForest)

# Define file path
importance_plot_path <- here::here("results", "figures", "rf_feature_importance.png")

# Save the plot
png(filename = importance_plot_path, width = 800, height = 600)
varImpPlot(rf_final_model)
dev.off()

```

We see that the model does not perform as well on the test data (AUC=0.56), and is able to distinguish the severity only 56% of the time.
The most important feature is the patient age.
However, this variable has imputations and thus may affect the dependability of the model.
Overall, the model performs decently on train data, but not on test data.
More data and variables can help improve the fit.
With this data, this model is not useful in distinguishing between serious and non-serious cases.

Let's create a summary table for all the models:
```{r}
library(pROC)
library(caret)

# Logistic Regression Model
logit_preds <- predict(logit_model, newdata = train_data)
logit_acc <- confusionMatrix(logit_preds, train_data$serious_bin)$overall['Accuracy']
logit_auc <- roc(train_data$serious_bin, predict(logit_model, newdata = train_data, type = "prob")[,2])$auc
logit_sens <- confusionMatrix(logit_preds, train_data$serious_bin)$byClass['Sensitivity']
logit_spec <- confusionMatrix(logit_preds, train_data$serious_bin)$byClass['Specificity']

# Decision Tree Model (now using dt_model)
dt_preds <- predict(dt_model, newdata = train_data)
dt_acc <- confusionMatrix(dt_preds, train_data$serious_bin)$overall['Accuracy']
dt_auc <- roc(train_data$serious_bin, predict(dt_model, newdata = train_data, type = "prob")[,2])$auc
dt_sens <- confusionMatrix(dt_preds, train_data$serious_bin)$byClass['Sensitivity']
dt_spec <- confusionMatrix(dt_preds, train_data$serious_bin)$byClass['Specificity']

# Random Forest Model
rf_preds <- predict(rf_model, newdata = train_data)
rf_acc <- confusionMatrix(rf_preds, train_data$serious_bin)$overall['Accuracy']
rf_auc <- roc(train_data$serious_bin, predict(rf_model, newdata = train_data, type = "prob")[,2])$auc
rf_sens <- confusionMatrix(rf_preds, train_data$serious_bin)$byClass['Sensitivity']
rf_spec <- confusionMatrix(rf_preds, train_data$serious_bin)$byClass['Specificity']

# Create summary data frame with results
model_summary <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Accuracy = c(logit_acc, dt_acc, rf_acc),
  AUC = c(logit_auc, dt_auc, rf_auc),
  Sensitivity = c(logit_sens, dt_sens, rf_sens),
  Specificity = c(logit_spec, dt_spec, rf_spec)
)

# Convert to huxtable
model_table <- as_hux(model_summary)

# Format the table
model_table <- set_bold(model_table, 1, 1:ncol(model_table))  # Make the entire header row bold
model_table <- set_align(model_table, 1, 1:ncol(model_table), "center")  # Center-align the entire header row
model_table <- set_all_borders(model_table, 1)

# Print the table
model_table

# Save the table
model_table_path = here("results", "tables", "model_table.rds")
saveRDS(model_table, file = model_table_path)
```
We see that sensitivity is 0 and specificity is high, meaning all the models always predict 'serious'.
Such problems arise due to class imbalances, and may require advanced methods to deal with.